---
title: "Homework 1"
author: "jys2137"
date: 2021-09-29
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

The purpose of this file is to present the answers to Homework 1, an assignment reinforcing ideas in the **building blocks** topic for P8105. 


## Problem 0.1

This “problem” focuses on the use of R Markdown to write reproducible reports, GitHub for version control, and R Projects to organize the work.

To that end:
  
* I created a public GitHub repo + local R Project; naming this repo / directory p8105_hw1_jys2137
* I created a single .Rmd file named p8105_hw1_jys2137.Rmd that renders to github_document
* I submitted a link to my repo via Courseworks


## Problem 0.2

This “problem” focuses on correct styling for solutions to Problems 1 and 2. 

To that end, I made sure to:
  
* use meaningful variable / object names
* use readable code (one command per line; adequate whitespace and indentation; etc)
* use clearly-written text to explain code and results
* avoid using superfluous code (no unused variables are defined; no extra library calls; etc)


## Problem 1

### Creating the data frame

The code chunk below creates a **dataframe** comprised of:
  
* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor "levels"

```{r df_create}
problem1_df = 
  tibble(
    samp = rnorm(10),
    samp_pos = samp > 0,
    vec_char = c("monday", "tuesday", "wednesday", "thursday", "friday", "saturday", "sunday", "someday", "birthday", "yesterday"),
    vec_factor = factor(c("sunny", "sunny", "sunny", "cloudy", "cloudy", "cloudy", "rainy", "rainy", "rainy", "rainy"))
)
```

### Taking the mean of each variable in the data frame

Here, we try to take the means of the variables `samp`, `samp_pos`, `vec_char`, and `vec_factor`.

```{r mean_pull_var}
mean(pull(problem1_df, samp))
mean(pull(problem1_df, samp_pos))
mean(pull(problem1_df, vec_char))
mean(pull(problem1_df, vec_factor))
```

Note that when trying to take the mean, the following **works**:
  
* The variable `samp`. _This is numeric._ 
* The variable `samp_pos`. _This is logical._

When trying to take the mean, the following **does not work**:

* The variable `vec_char`. _This is neither numeric nor logical. We get a warning message and a subsequent output of `NA`._
* The variable `vec_factor`. _As above, this is neither numeric nor logical. We get message and a warning and a subsequent output of `NA`._

### Explicitly converting variables to numeric

The chunk below applies the `as.numeric` function to the logical, character, and factor variables.

```{r try_numeric, warning = FALSE, results = "hide", message = FALSE}

as.numeric(pull(problem1_df, samp))
as.numeric(pull(problem1_df, samp_pos))
as.numeric(pull(problem1_df, vec_char))
as.numeric(pull(problem1_df, vec_factor))
```

* Note: the code for `samp` works because it is numeric.
* Converting `samp_pos` to numeric **works**. This happens because it considered "FALSE" as 0 and "TRUE" as 1, which helps explain why it works when we try to take the mean since it can be calculated as a fraction of the values that are "TRUE".
* Converting `vec_char` to numeric **does not work**. We get an output of NAs because the values cannot be interpreted as numeric, which helps exlain why we could not take the mean.
* Converting `vec_factor` to numeric **works**. This happens because it considered `"sunny"` as 3, `"cloudy"` as 1, and `"rainy"` as 2, which explains why we were able to take the mean.

